---
title: "Case Study"
output: case_study_notebook
---

Load in the important libraries.

```{r}
library(tidyverse)
library(dplyr)
library(tidyr)
library(ggplot2)
```

## **Problem 1 – Data handling, analysis and plotting**	

The first problem of the case study builds on the data in the files p01-02_portfolio.csv and
p01-02_rates.csv. One file contains membership information for a Group Life portfolio and one
has information on the rates which should be charged.

```{r}
# Load the CSV file
portfolio_data <- read_delim(
  "Case Study/data/p01-02_portfolio.csv",
  delim = ";",
  show_col_types = FALSE
)

# View the data
print(portfolio_data)
```

```{r}
# Load the CSV file
rates_data <- read_delim(
  "Case Study/data/p01-02_rates.csv",
  delim = ";",
  show_col_types = FALSE
)

# View the data
print(rates_data)
```

### **Question a.**			

Read the data from the two files into R’s memory. The rates are applicable to each individual in the portfolio, depending on that individual’s age and gender. Combine the two datasets into a single table by looking up the rate for each line of the portfolio.

```{r}

```
